\Timestamp{2015-01-08}

\begin{df}[Vertex Cover]
	For a given graph $G(V, E)$ a \emphdef{vertex cover} ist a subset $C \subset V$, such that
	\begin{math}
		\forall e = \Set{u, v} \in E : C \cap e \neq \emptyset
	\end{math}
\end{df}

\begin{df}[Vertex Cover Problem]
	Given a graph $G(V, E)$, the vertex cover problem is to determine a vertex cover $C \subset V$, such that $|C|$ is minimal.
	\begin{note}
		\begin{itemize}
			\item
				The vertex cover problem is NP-hard.
			\item
				The vertex cover problem is a special case of the set cover problem.
				To see this, let the universe be the set of all edges $U := E$ and the subsets
				\begin{math}
					S_v := \Set{e \in E : v \in e} & v \in V,
				\end{math}
				i.e. sets of all edges adjacent to vertices $v \in V$.
		\end{itemize}
	\end{note}
\end{df}

\begin{alg}[Set Cover, Idea] \label{alg:set_cover_lp_rounded}
	\begin{algorithmic}
		\State{solve $LP$-relaxation to optimality}
		\State{choose all sets $S$ with $x_S > 0$}
	\end{algorithmic}
	This obviously yields a feasible solution.
\end{alg}

\begin{df}[ILP for vertex cover problem]
	Solve
	\begin{math}
		\min \sum_{v \in V} x_V
		\st
		\forall e = \Set{v, w} \in E : x_v + x_w \ge 1 \land x_v \in \Set{0, 1}.
	\end{math}
	Analogously the LP-relaxation is defined with $x_v \ge 0$.
\end{df}

\begin{lem}[Half integrality of vertex cover problem solution]
	The LP-relaxation of the vertex cover problem yields a solution with
	\begin{math}
		x_S \in \Set{0, \frac{1}{2}, 1}
	\end{math}
	for all sets $S$.
	\begin{proof}
		Show that every corner of the polyhedron of the vertex cover LP has coordinates $0, \frac{1}{2}, 1$.
		Assume $x \in \Q^n, n = |V|$ is a feasible non-half-integral solution to the LP.
		Define
		\begin{math}
			V_+ &:= \Set{v \in V & \frac{1}{2} < x_v < 1} &
			V_- &:= \Set{v \in V & 0 < x_v < \frac{1}{2}}
		\end{math}
		For $\eps > 0$ let
		\begin{math}
			y_v &:= \begin{cases}
				x_v + \eps & v \in V_+ \\
				x_v - \eps & v \in V_- \\
				x_v & \text{otherwise}
			\end{cases}, &
			z_v &:= \begin{cases}
				x_v - \eps & v \in V_+ \\
				x_v + \eps & v \in V_- \\
				x_v & \text{otherwise}
			\end{cases}.
		\end{math}
		Clearly $x_v, y_v, z_v$ are distinct and $x_v = \frac{1}{2} (y_v + z_v)$.

		If $y_v, z_v$ are feasible, we have a contradiction to $x_v$ being a corner.

		So consider an edge constraint $x_v + x_w \ge 1$ in the case, that $x_v + x_w = 1$ (else $\eps > 0$ could have been chosen small enough for $y_v, z_v$ to be feasible).
		The following possibilities arise
		\begin{enumerate}[i)]
			\item
				$x_v = 1 \land x_w = 0$,
			\item
				$x_v = 0 \land x_w = 1$,
			\item
				$x_v = x_w = \frac{1}{2}$,
			\item
				$v \in V_+, w \in V_-$,
			\item
				$v \in V_-, w \in V_+$.
		\end{enumerate}
		in the cases i) to iii) we have $y_v = z_v = x_v, y_w = z_w = x_w$.
		in the cases iv) and v) $y_v + y_w = x_v + x_w = z_v + z_w$.
		As such $x$ is feasible.
	\end{proof}
\end{lem}

\begin{kor}
	Algorithm \ref{alg:set_cover_lp_rounded} applied to the set cover problem induced by the vertex cover problem yields a solution at most double the optimal solution.
	\begin{note}
		More generally for any covering LP a bounded fractionality of $x_i \ge \frac{1}{f}$ immidiately yields a $f$-approximation.
	\end{note}
\end{kor}


